---
title: "Pseudo_Facebook_data"
output: html_document
---
```{r}
#install.packages('ggplot2')
library(ggplot2)
#install.packages('gridExtra')
library(gridExtra)
#install.packages('dplyr')
library(dplyr)
#install.packages('reshape2')
library(reshape2)
```

Read in the psuedo Facebook data.
```{r}
#use sep to indicate file is tab seperated
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```

## Friend count
Let's see look in to the friend count and see if there's much of a difference in friend count between male and female users. We would like to omit the rows which have 'NA' as their gender by subseting the data to only include the rows that don't have NA as gender.
```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
From this plot it looks like male users have a higher proportion of low friend count 
than female useers. Let's look at the statistics of friend count by gender to 
get a clearer idea of the difference in friend count.
```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```
Both the median and mean friend count for female users is higher than for males. 
The median friend count will be a more robust statistic in this case as extreme 
outliers may skew the mean friend count. 

Visulaising this with box plots. Adjust the code to focus on users who have 
friend counts between 0 and 1000.
```{r}
ggplot(aes(x = gender, y=friend_count), 
       data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,1000))
```

### Proportion of friend count
Comparing male and female friend counts in an overlay with a *frequency polygon* 
rather than two histograms side by side. 
Change the y axis to show prportions instead of counts.

```{r Frequency Polygons}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```
A higher proportion of male users have a low friend count than women do. Towards 
the higher end of friend count (300+) male and female users seem to be close.

Let's look in to some of the factors that might affect friend count.

### Tenure
Friend count will be affected by how long a user has been a member of Facebook so let's plot the tenure for Facebook users.
```{r Tenure Histogram by Year}
qplot(x=tenure/365, data = pf, binwidth=0.25, 
      xlab ='Number of years using Facebook',
      ylab='Number of users in sample',
      color=I('black'), fill=I('#F79420') )
```
The majority of users have been on Facebook for less than 2.5 years. However, 
those that have been on Facebook longer will likely have gained more friends over
time than the newer users. 


## User Ages
Let's see what age groups use Facebook. 
```{r User Ages}
summary(pf$age) #give the min and max ages help determine scale for x axis
qplot(x=age, data=pf, binwidth=1, xlab = "age", 
      ylab='Number of users in sample',
      color=I('black'), fill=I('#5760AB')) +
      scale_x_continuous(breaks = seq(13,113,5)) #seq(0,113,5) goes from 0 to 113 in units of 5 on x axis
```
The summary shows the minimum age is 13 and the max is 113. This is useful as you 
can't use facebook if you're younger than 13 and we're not confident about the 
data we have for the over 90s are truthful. 

Now let's see if friend count varies by the age of users on Facebook.
```{r Scatterplots}
ggplot(aes(x=age, y=friend_count), data =pf) + geom_jitter(alpha=1/20) + 
  xlim(13, 90)

summary(pf$age)
```
* There are vertical bars where people have lied about their age - e.g. 69 and 
100. These users are likely to be teenagers or fake accounts given the huge 
friend counts.
* The users with the most friends are typically under 30.
* The bulk of young users have friend count below 1000.

Transform the plot using square root:
```{r Coord_trans()}
ggplot(aes(x=age, y=friend_count), data =pf) + 
  geom_point(alpha=1/20, position = position_jitter(h=0)) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt') #transform y axis to square root values
```
With this plot it's much easier to see the distribution of friend count 
conditional on age. For example we can see threshold count above which there are 
very few users.

We can split up the datframe and apply a function (to find the mean and median)
to some parts of the data. It can be helpful to not just plot all points but to
plot against the mean or median friend count for each age as a refernce.
```{r Conditional Means}
#group dataframe by age and store it in age_groups
age_groups <- group_by(pf, age)

# summarise this new grouping of data and create new variables of mean
# friend count, median friend count & the number of people in each group
# Save all of this in a new variable pf.fc_by_age
pf.fc_by_age <- summarise(age_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) #number of users in each group. 
# n fxn can only be used in summarise and it reports how many people are in each group

#Order data by age
pf.fc_by_age <- arrange(pf.fc_by_age, age)

#print first few rows of dataframe
head(pf.fc_by_age)
```

Overlay plots of the mean and the 10th, 90th and 50th percent quartile in order to give some more context to the graph.

```{r Overlaying Summaries with Raw Data}
library(ggplot2)
ggplot(aes(x=age, y= friendships_initiated), data = pf) +
  geom_point(alpha = 1/10, position = position_jitter(h=0), color='orange') +
  coord_cartesian(xlim = c(13, 90), ylim = c(0, 1000)) +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat= 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype =2, color='blue') +
  geom_line(stat= 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype =2, color='blue') +
  geom_line(stat= 'summary', fun.y = quantile, fun.args = list(probs = .5), color='blue')
# fun.args = list(probs = .9) gives the 90th percent quantile, linetype=2 makes it dashed 
```

***
### Correlation between age and friend count
If the correlation coefficient is close to 1, it would indicate that the 
variables are positively linearly related and the scatter plot falls almost 
along a straight line with positive slope. A rule of thumb is that a correlation 
greater than 0.3 or less than -0.3 is meaningful, but small. around 0.5 is 
moderatre and 0.7 is large.
```{r Correlation}
#the default method for computing the correlation coefficient is Pearson. You do not need to pass the method parameter when calculating the Pearson Product Moment Correlation.
cor.test(pf$age, pf$friend_count)

#Another method:
# The with fxn lets us evaluate an R expression in an environment constructed from the data. 
# This codde gives the same result as above
with(pf, cor.test(age, friend_count, method='pearson'))
```
The correlation between age and friend count: -0.027 . This implies that there 
isn't a meaningful relationship between age and friend count.

### Correlation on Subsets
Based on the correlation co efficient and the above plot we just observed that 
the relationship between age and friend count is not linear. It isn't monotonic, 
either increasing or decreasing. Furthermore, based on the plot we know that we 
maybe don't want to include the older ages in our correlation number since older 
ages are likely to be incorrect. 
Lets redo the plot for users aged 70 or less: 
```{r Correlation on Subsets}
with(subset(pf, age<=70), cor.test(age, friend_count))
```
This tells a different story about a negative relationship between age and 
friend count. As age increases friend count decreases. The pearson product 
measures the strength of a relationship between any two variables. But there can
be lots of other types of relationships. 

***

### Comparing age, gender and friend count
Boxplot of ages by gender. Here we can see the averages by a shape marked 'x'. 
Since male users are a bit younger we might think a simple male to female 
comparison doesn't capture the substantial differences in friend count.  
```{r}
ggplot(aes(x=gender, y=age), data= subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y=mean, geom = 'point', shape=4)
```

Lets make a graph of median friend count:
```{r}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line( aes(color=gender), stat='summary', fun.y=median)
```
We can see that nearly everywhere the median friend count is larger for women 
than it is for men. There are some exceptions which include the noise around the 
old users. We're not really confident about these reported older ages. Users 
reporting age 70 seem to more or less have the same friend count regardless 
of gender. 

Write code to create a new data frame that contains information on each age AND 
gender group.
The data frame should contain the following variables:
* mean_friend_count
* median_friend_count
* n (the number of users in each age and gender grouping)
```{r Third Qualitative Variable}
new_groupings <- group_by(pf, age, gender)
pf.fc_by_age_gender <- summarise(new_groupings,
                                 mean_friend_count = mean(friend_count),
                                 median_friend_count = median(friend_count),
                                 n=n())
head(pf.fc_by_age_gender)
```

```{r Plotting Conditional Summaries}
ggplot(aes(x=age, y=median_friend_count), data = subset(pf.fc_by_age_gender, !is.na(gender))) +
  geom_line(aes(color=gender))
```
It seems like the largest difference between friend count for male and female
occurs between young users. 

***

### Thinking in Ratios
Let's answer the question how many more times friends does the average female 
user have than the male user. To answer that question we need to reshape our 
data. Right now the data is in long format. We have many rows and the variables 
that we grouped over, male and female, have been repeated for each year. 
Let's convert from long format to wide format. This new dataframe will have one 
row for each age and then we'll put the median friend count inside of rows for 
males and females. 

```{r}
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)

```

## Putting it all together - Tenure & Age
For friend count we need to take in to account how long people have been on 
Facebook as someone who's been using it for a long time will likely have gained 
more friends. This data was collected in 2014 so we're going to take that in to 
account.
Create a variable called year_joined in the data frame using the variable tenure 
and 2014 as the reference year. 
Tenure variable = how many days since registering with Facebook. 
```{r Third Quantitative Variable}
pf$year_joined <- floor(2014-pf$tenure/365) 
# the floor() function to round down to the nearest integer
```

Let's  look at a summary of the data on year joined
```{r}
summary(pf$year_joined)
#table format of year joined data
table(pf$year_joined)
```
There's not much info on the earlier years - 2005 & 2006. To increase the info
we have in each tenure category we can group some of this data together. We're 
going to create 4 bins for the years. The bins will be (i) 2004-2009 
(ii) 2009 - 2011 (iii) 2011 - 2012 (iv) 2012 - 2014
```{r Cut a Variable}
pf$year_joined_bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
table(pf$year_joined_bucket)
```

We can graph friend_count vs. age so that each year_joined.bucket is a line 
tracking the median user friend_count across age. To put these cohorts in to 
prespective we can plot the grand mean on here as well. The grand mean is the 
overall mean of friend count vs age.
```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined_bucket))) +
  geom_line(aes(color = year_joined_bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype=2)
```
We can see that users with longer tenure tend to have  higher friend counts with 
the exception of older users of about 80+. 
We might increase our confidence that this observation isn't just an artefact of 
the time users have had to accumulate friends. 

## Friending rate

We could also look at tenure and friend count as a rate instead. 
For example we could see how many friends does a user have for each day since 
using the service. 
Subset the data so you only consider users that have at least one day of tenure
```{r Friending Rate}
with(subset(pf, tenure>=1), summary(friend_count/tenure))
```

Do new users go on  a friending spree? Or do users with greater tenure 
friend more people? Plot friendships initiated per day vs. tenure. 
```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure),data = subset(pf, tenure>=1)) +
  geom_line(aes(color = year_joined_bucket), stat = "summary", fun.y = mean) 
```
It looks like users with more tenure typically initiate less friendships.

There's a lot of noise in our graph since we're plotting the mean of y for every
possible tenure x value. We can reduce this noise by increasing the bin width on 
the x axis. 

```{r Bias-Variance Tradeoff Revisited}

p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined_bucket),
            stat = 'summary',
            fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined_bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined_bucket),
            stat = "summary",
            fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined_bucket),
            stat = "summary",
            fun.y = mean)
grid.arrange(p1, p2, p3, p4, ncol=1)

```













